PANCHANGE_analysis

### 2020-12-09 PANCHANGE_analysis.Rmd
moved functions for forest out of script

files.sources = list.files("FunctionLib/")
sapply(files.sources, source)

@@

###Function to create formula

```{r function create formula for lin models}

model1b.lm.form <- function(outcome, explan_var, data_set) {

  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c(explan_var,
                             "depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23"
  )
  # create the formula
  glm.formula <- paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ ")

  # assign it globally with the right variable name.
  assign(paste0(outcome,".", explan_var, ".glmformula"),
         glm.formula,
         envir = .GlobalEnv

  )
}

@@

## Function to create dataframes ready for forest plots from lm models. 
```{r function to create data for lm forest plots}

# function to create dataframe from model output

lmoutput_format <- function(lmobject,explan_var, group, dfname){

  ## Row labels
  independent.variables <- c("intercept",
                             explan_var,
                             "comorbid depression and anxiety compared to control",
                             "depressive disorder compared to control",
                             "anxiety disorder compared to control",
                             "eating disorders compared to control",
                             "obsessive compulsive disorders compared to control",
                             "comorbid psychotic and bipolar disorder compared to control",
                             "psychotic disorder compared to control",
                             "bipolar disorder compared to control",
                             "mhd.posttraumatic_stress_disorder_ptsd compared to control",
                             "autism_spectrum_disorder compared to control",
                             "mhd.attention_deficit_hyperactivity_disorder compared to control",
                             "mhd.personality_disorder compared to control",
                             "Female compared to male",
                             "Non-binary/Self-defined compared to male",
                             "16-18 compared to 26-35",
                             "19-25 compared to 26-35",
                             "36-45 compared to 26-35",
                             "46-55 compared to 26-35",
                             "56-65 compared to 26-35",
                             "66-70 compared to 26-35",
                             "71-75 compared to 26-35",
                             "76+ compared to 26-35",
                             "Mixed or multiple ethnicities compared to White",
                             "Asian compared to White",
                             "Black compared to White",
                             "Other compared to White",
                             "key_worker compared to non-keyworker",
                             "time_diff_coping_lockdown_march_23")
  ## categories to facet on  
  categories <- c("Intercept",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Pre-existing mental health",
                  "Gender",
                  "Gender",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Age",
                  "Ethnicity",
                  "Ethnicity",
                  "Ethnicity",
                  "Ethnicity",
                  "Other",
                  "Other")

  ## Disorder dimension  
  groups <- c(group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group,
                  group)

  # create data frame
  df <- as.data.frame(coef(summary(lmobject)),row.names=F)

  # add row labels
  df$RowLabels <- independent.variables

  # Add category labels
  df$Categories <- categories

  # Add group labels
  df$Group <- groups

  # calculate upper and lower confidence intervals
  df <- df %>%
    mutate(LowerBound = Estimate - (1.96*`Std. Error`),
           UpperBound = Estimate + (1.96*`Std. Error`))

  # drop intercept row
    df <- df[(df$RowLabels != 'intercept'),]

  # order factors

  # assign dataframe to global environment
  assign(dfname, df, envir = knitr::knit_global())

## Function to create forest plots

```{r function for creating forest plots}

lm.forest <- function(data,title){

  p = ggplot(data=data,
             aes(x =RowLabels,y = Estimate,
                colour = Group)) +

    geom_point(size = 2,
               position = position_dodge(0.9)) +

    geom_hline(aes(fill=Group),
               yintercept =0, 
               linetype=2)  +

    xlab('Group') + 
    ylab("Estimate (95% Confidence Interval)")  +
  ggtitle(title) +

    geom_errorbar(aes(ymin = LowerBound,
                      ymax = UpperBound),
                  width = .5,
                  position = position_dodge(0.9))  +

    facet_wrap(. ~ Categories,
               strip.position="left",nrow=5,scales = "free_y") +

  scale_colour_manual(values=COPINGpalette5) +

    theme(plot.title=element_text(size=20,face="bold"),
          axis.text.y=element_text(face="bold"),
          axis.text.x=element_text(face="bold"),
          axis.title.x=element_text(size=14,face="bold"),
          axis.title.y=element_blank(),
          strip.text.y = element_text(angle=180,face="bold"),
          panel.background = element_blank(),
          strip.background = element_rect(fill = "light grey"),
          panel.border = element_rect(colour = "grey",fill = NA)) +
    coord_flip()
  return(p)

@@

## Forest plots
### disorder
```{r forest plot by disorder prepan, fig.height=20, fig.width=12}

@@

```{r forest plot by disorder retro, fig.height=20, fig.width=12}

@@

### disorder (no PCL)
```{r forest plot by disorder prepan no pcl, fig.height=20, fig.width=12}

@@

```{r forest plot by disorder retro no pcl, fig.height=20, fig.width=12}

@@

### time point 
```{r forest plot by time point gad, fig.height=20, fig.width=12}

@@

```{r forest plot by time point phq, fig.height=20, fig.width=12}

@@

```{r forest plot by time point pcl, fig.height=14, fig.width=8}

@@

```{r forest plot by sample retro only gad, fig.height=20, fig.width=12}

@@

```{r forest plot by sample retro only phq, fig.height=20, fig.width=12}

@@

```{r forest plot by sample retro only pcl, fig.height=20, fig.width=12}

>>>>

files.sources = paste0("FunctionLib/",list.files("FunctionLib/"))

sapply(files.sources, source)

@@

@@

@@

## Forest plots
### disorder
```{r forest plot by disorder prepan, fig.height=20, fig.width=20, dpi=120}

@@

```{r forest plot by disorder retro, fig.height=20, fig.width=20, dpi=120}

@@

### disorder (no PCL)
```{r forest plot by disorder prepan no pcl, fig.height=20, fig.width=20, dpi=120}

@@

```{r forest plot by disorder retro no pcl, fig.height=20, fig.width=20, dpi=120}

@@

### time point 
```{r forest plot by time point gad, fig.height=20, fig.width=20, dpi=120}

@@

```{r forest plot by time point phq, fig.height=20, fig.width=20, dpi=120}

@@

```{r forest plot by time point pcl, fig.height=20, fig.width=20, dpi=120}

@@

```{r forest plot by sample retro only gad, fig.height=20, fig.width=20, dpi=120}

@@

```{r forest plot by sample retro only phq, fig.height=20, fig.width=20, dpi=120}

@@

```{r forest plot by sample retro only pcl, fig.height=20, fig.width=20, dpi=120}
